---
title: "AmpBS-Pipeline"
author: "Joschka Hey & Maximilian Schönung"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r Set-Up, include=FALSE}
# run Rscript -e "rmarkdown::render('pipeline_bisSeq.Rmd')" ~/Documents/Data/AmpBS_JMML_tech_val/cov_files/ "~/Desktop/sites_analysis.bed" 10 "~/Desktop/meta.txt" "~/Desktop/"

### Set-Up ###

# Load the libraries
library(BiSeq)
library(reshape2)
library(ggplot2)
library(GenomicRanges)
library(pheatmap)
library(FactoMineR)

# Load the arguments
args <- commandArgs(T)

# Specify the paths based on the arguments
#args <- c("~/Documents/Data/AmpBS_JMML_tech_val/cov_files/","~/Desktop/sites_analysis.bed",10,"~/Desktop/meta.txt")

file <- paste0(args[1],list.files(args[1]))
file_ind <- stringr::str_detect(file,".cov")
file <- file[file_ind] 
cg_in <- read.delim(args[2],header=F)
cutoff <- as.numeric(args[3])
meta <- read.delim(args[4],header=T,stringsAsFactors = F)
output.dir <- args[5]

#meta[,2] <- c(rep("a",10),rep("b",10),rep("c",12))

# shape the meta-data
rownames(meta) <- meta$UPN

# Load the AmpBS-Seq data in an summarized experiment
meth.set <- readBismark(file,colData=meta)
```

# Quality Control

#### Total Reads
```{r echo=FALSE}
total.reads <- totalReads(meth.set)
total.reads <- data.frame("Sample"=colnames(total.reads),"Sites"=colSums(total.reads))

ggplot(total.reads,aes(Sample,log10(Sites)))+
  geom_bar(stat="identity")+
  theme_minimal()+
  ylab("log10(Total Reads per Sample)")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

In total `r log10(sum(total.reads$Sites))` x 10^5 reads were aligned and detected. 

<br>
<br>

#### Detected CpGs

```{r echo=FALSE}
cg.cov <- covStatistics(meth.set)
cg.cov <- data.frame("Sample"=names(cg.cov$Covered_CpG_sites),"CpG"=cg.cov$Covered_CpG_sites)

ggplot(cg.cov,aes(Sample,CpG))+
  geom_bar(stat="identity")+
  theme_minimal()+
  ylab("Covered CpG sites")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

<br>
<br>


#### Numer of reads per sample versus detected CpG sites

```{r echo=FALSE}
reads_cg <- cbind(total.reads,"CpG"=cg.cov$CpG)

ggplot(reads_cg,aes(log10(Sites),CpG))+
  geom_point()+
  theme_minimal()+
  geom_smooth(method="lm")
```

<br>
<br>

#### Coverage per CpG

```{r echo=FALSE}
cov_cg <- totalReads(meth.set)
cov_cg.melt <- melt(cov_cg)
cov_cg.melt <- cov_cg.melt[cov_cg.melt$value>0,]

ggplot(cov_cg.melt,aes(Var2,log10(value)))+
  geom_boxplot()+
  theme_minimal()+
  xlab("Sample")+
  ylab("log10(Coverage per CpG)")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

#### Covergae per CpG after coverage filtering

```{r echo=FALSE}
meth.clean <- BiSeq::filterByCov(meth.set,as.numeric(cutoff),global=FALSE)

cov_cg <- totalReads(meth.clean)
cov_cg.melt <- melt(cov_cg)
cov_cg.melt <- cov_cg.melt[cov_cg.melt$value>0,]

ggplot(cov_cg.melt,aes(Var2,log10(value)))+
  geom_boxplot()+
  theme_minimal()+
  xlab("Sample")+
  ylab("log10(Coverage per CpG)")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

<br>
<br>

#### Detected CpGs after coverage filtering

```{r echo=FALSE}
cg.cov <- covStatistics(meth.clean)
cg.cov <- data.frame("Sample"=names(cg.cov$Covered_CpG_sites),"CpG"=cg.cov$Covered_CpG_sites)

ggplot(cg.cov,aes(Sample,CpG))+
  geom_bar(stat="identity")+
  theme_minimal()+
  ylab("Covered CpG sites")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

<br>
<br>

#### Subset with the CpGs of interest

```{r echo=FALSE}
cg_in.ranges <- GRanges(seqnames = cg_in$V1, ranges = IRanges(cg_in$V2,width=cg_in$V3-cg_in$V2+1) ,cg_is=cg_in$V4)

meth.sub <- subsetByOverlaps(meth.clean,cg_in.ranges)
```


#### Covergae per CpG of interest after coverage filtering

```{r echo=FALSE}

cov_cg <- totalReads(meth.sub)
cov_cg.melt <- melt(cov_cg)
cov_cg.melt <- cov_cg.melt[cov_cg.melt$value>0,]

ggplot(cov_cg.melt,aes(Var2,log10(value)))+
  geom_boxplot()+
  theme_minimal()+
  xlab("Sample")+
  ylab("log10(Coverage per CpG)")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

<br>
<br>

#### Detected CpG of interest after coverage filtering

```{r echo=FALSE}
cg.cov <- covStatistics(meth.sub)
cg.cov <- data.frame("Sample"=names(cg.cov$Covered_CpG_sites),"CpG"=cg.cov$Covered_CpG_sites)

ggplot(cg.cov,aes(Sample,CpG))+
  geom_bar(stat="identity")+
  theme_minimal()+
  ylab("Covered CpG sites")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

<br>
<br>

# PCA

#### PCA with all CpGs after coverage filtering

```{r echo=FALSE}
# make a rel Bis-Object
meth.clean.rel <- rawToRel(filterByCov(meth.set,as.numeric(cutoff),T))
meth.cluster <- assay(meth.clean.rel)
pca <- PCA(t(meth.cluster),graph = F,scale.unit = F)
pca.df <- pca$ind$coord
colnames(pca.df) <- paste0("PC",1:ncol(pca.df))


ggplot(as.data.frame(pca.df),aes(PC1,PC2))+
  geom_point()+
  theme_minimal()+
  xlab(paste0("PC1 (",round(pca$eig[1,2],2),"% Variance)"))+
  ylab(paste0("PC2 (",round(pca$eig[2,2],2),"% Variance)"))

#Color with pheno data
if(ncol(meta)>1){
  for(i in 2:ncol(meta)){
    coloring <- factor(meta[,i])
    
  p <- ggplot(as.data.frame(pca.df),aes(PC1,PC2,color=coloring))+
  geom_point()+
  theme_minimal()+
  xlab(paste0("PC1 (",round(pca$eig[1,2],2),"% Variance)"))+
  ylab(paste0("PC2 (",round(pca$eig[2,2],2),"% Variance)"))
  print(p)  
  }
}

```


<br>
<br>


#### PCA with CpGs of interest after coverage filtering

```{r echo=FALSE}
# make a rel Bis-Object
meth.clean.rel <- rawToRel(filterByCov(meth.sub,as.numeric(cutoff),T))
meth.cluster <- assay(meth.clean.rel)
pca <- PCA(t(meth.cluster),graph = F,scale.unit = F)
pca.df <- pca$ind$coord
colnames(pca.df) <- paste0("PC",1:ncol(pca.df))


ggplot(as.data.frame(pca.df),aes(PC1,PC2))+
  geom_point()+
  theme_minimal()+
  xlab(paste0("PC1 (",round(pca$eig[1,2],2),"% Variance)"))+
  ylab(paste0("PC2 (",round(pca$eig[2,2],2),"% Variance)"))

#Color with pheno data
if(ncol(meta)>1){
  for(i in 2:ncol(meta)){
    coloring <- factor(meta[,i])
    
  p <- ggplot(as.data.frame(pca.df),aes(PC1,PC2,color=coloring))+
  geom_point()+
  theme_minimal()+
  xlab(paste0("PC1 (",round(pca$eig[1,2],2),"% Variance)"))+
  ylab(paste0("PC2 (",round(pca$eig[2,2],2),"% Variance)"))
  print(p)  
  }
}

```

# Clustering

<br>

All Clusterings are performed with manhattan distance and ward.D2 clustering.

#### Unsupervised hierachical clustering with all CpGs after coverage filtering

```{r echo=FALSE}
# make a rel Bis-Object
meth.clean.rel <- rawToRel(filterByCov(meth.set,as.numeric(cutoff),T))
meth.cluster <- assay(meth.clean.rel)


pheatmap(meth.cluster,
         clustering_distance_cols = "manhattan",
         clustering_method = "ward.D2",
         show_rownames = F)

#Color with pheno data
if(ncol(meta)>1){
  for(i in 2:ncol(meta)){
    anno_df <- data.frame(meta[,2:ncol(meta)])
    rownames(anno_df) <- meta$UPN
    colnames(anno_df) <- colnames(meta)[2:ncol(meta)]
    
    p <- pheatmap(meth.cluster,
         clustering_distance_cols = "manhattan",
         clustering_method = "ward.D2",
         show_rownames = F,
         annotation_col = anno_df)
    print(p)
    
  }}


```

<br>
<br>

#### Unsupervised hierachical clustering with CpGs of interest after coverage filtering

```{r echo=FALSE}
# make a rel Bis-Object
meth.clean.rel <- rawToRel(filterByCov(meth.sub,as.numeric(cutoff),T))
meth.cluster <- assay(meth.clean.rel)


pheatmap(meth.cluster,
         clustering_distance_cols = "manhattan",
         clustering_method = "ward.D2",
         show_rownames = F)


#Color with pheno data
if(ncol(meta)>1){
  for(i in 2:ncol(meta)){
    anno_df <- data.frame(meta[,2:ncol(meta)])
    rownames(anno_df) <- meta$UPN
    colnames(anno_df) <- colnames(meta)[2:ncol(meta)]
    
    p <- pheatmap(meth.cluster,
         clustering_distance_cols = "manhattan",
         clustering_method = "ward.D2",
         show_rownames = F,
         annotation_col = anno_df)
    print(p)
    
  }} 
```


<br>
<br>


```{r echo=FALSE}
### save the files

## Cut-off filtering
meth.clean.rel <- rawToRel(filterByCov(meth.set,as.numeric(cutoff),T))
meth.table <- assay(meth.clean.rel)
ids <- as.data.frame(rowRanges(meth.clean.rel))
rownames(meth.table) <- paste0(ids$seqnames,":",ids$start)
meth.table <- round(meth.table,2)
write.table(meth.table,paste0(args[5],Sys.Date(),"_beta_values_cutoff_",args[3],".txt"),sep="\t",quote=F)

coverage.df <- totalReads(filterByCov(meth.set,as.numeric(cutoff),T))
rownames(coverage.df) <- paste0(ids$seqnames,":",ids$start)
write.table(coverage.df,paste0(args[5],Sys.Date(),"_coverage_cutoff_",args[3],".txt"),sep="\t",quote=F)

## CpG of interest
meth.clean.rel <- rawToRel(filterByCov(meth.sub,as.numeric(cutoff),T))
meth.table <- assay(meth.clean.rel)
ids <- as.data.frame(rowRanges(meth.clean.rel))
rownames(meth.table) <- paste0(ids$seqnames,":",ids$start)
meth.table <- round(meth.table,2)
write.table(meth.table,paste0(args[5],Sys.Date(),"_CpG_interest_beta_values_cutoff_",args[3],".txt"),sep="\t",quote=F)

coverage.df <- totalReads(filterByCov(meth.sub,as.numeric(cutoff),T))
rownames(coverage.df) <- paste0(ids$seqnames,":",ids$start)
write.table(coverage.df,paste0(args[5],Sys.Date(),"_CpG_interest_coverage_cutoff_",args[3],".txt"),sep="\t",quote=F)

```

