---
title: "Report AmpBS-Seq analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
params:
  set: NA
  set_filt: NA
  set_filtR: NA
  cov_co: NA
  bed_regions: NA
  meta_df: NA
---

```{r setup, include=FALSE}
data <- params$set
data.clean <- params$set_filt
data.cleanR <- params$set_filtR
co <- params$cov_co
bed <- params$bed_regions
meta <- params$meta_df
```

<br>

## Info metrices
<br>
The following table presents an overview of the uploaded regions, sample annotations and covered CpG sites:
<br>
<br>

```{r echo=FALSE}
info.tab <- data.frame("Paramter"=c("Number of Samples","Meta-Data","Sites (Total)","Sites (Coverage-Filtered)","Regions (bed-Input)"),
    "Value"=c(length(colnames(data)),
    stringr::str_flatten(colnames(data@colData),", "),
    nrow(data@elementMetadata),
    nrow(BiSeq::filterByCov(data,as.numeric(100),global=FALSE)@elementMetadata),
    nrow(bed@elementMetadata)))
DT::datatable(info.tab)
```

<br>

Additionally, the following table provides coverage information for the uploaded samples:
<br>
<br>

```{r echo=FALSE}
    cg_tab <- cbind("Covered CpG"=BiSeq::covStatistics(data)$Covered_CpG_sites,
                    "Median Coverage"=BiSeq::covStatistics(data)$Median_coverage)
    rownames(cg_tab) <- meta$UPN
    DT::datatable(cg_tab)
```

<br>
<br>

## Quality Control (QC) Plots
<br>
In this section, basic QC statistics are visualized. First we will consider all mapped reads from the Bismark coverage files.

<br>
<br>

#### Barplot of the total reads per sample (log10 scale):
```{r echo=FALSE}
qc.plots <- qc_plots(data)
plot(qc.plots[[1]])
```

#### Barplot showing the number of covered CpG sites per sample:
```{r echo=FALSE}
plot(qc.plots[[2]])
```

#### Correlation between coverage and detected CpG sites:
The number of total reads per sample (log10 scale) is plotted on the x-axis against the number of detected CpG sites (y-axis). 
```{r echo=FALSE}
plot(qc.plots[[3]])
```

#### Boxplot showing the number of reads per CpG site (log10 scale):
```{r echo=FALSE}
plot(qc.plots[[4]])
```

<br>
<br>

## Quality Control (QC) Plots after coverage filtering
<br>
Now let's inspect the data after filtering for coverage. Therefore, CpG sites with less reads than the selected coverage filter will be discarded.
Your current coverage filter is set to:
<br>
```{r echo=FALSE}
co
```

<br>
<br>

#### Barplot showing the number of covered CpG sites per sample:
```{r echo=FALSE}
qc.plots.filt <- qc_plots(data.clean)
plot(qc.plots.filt[[2]])
```

#### Correlation between coverage and detected CpG sites:
The number of total reads per sample (log10 scale) is plotted on the x-axis against the number of detected CpG sites (y-axis). 
```{r echo=FALSE}
plot(qc.plots.filt[[3]])
```

#### Boxplot showing the number of reads per CpG site (log10 scale):
```{r echo=FALSE}
plot(qc.plots.filt[[4]])
```

<br>
<br>


## Quality Control (QC) based on region
<br>
Let's focus furthermore on the selected input regions. You have chosen the following regions:
```{r echo=FALSE}
DT::datatable(as.data.frame(bed)[,1:4])
```

<br>
<br>

#### Barplot showing the number of covered CpG sites in the chosen regions:
```{r echo=FALSE}
qc.plots.filtR <- qc_plots(data.cleanR)
plot(qc.plots.filtR[[2]])
```

#### Boxplot showing the number of reads per CpG site for the chosen regions (log10 scale):
```{r echo=FALSE}
plot(qc.plots.filtR[[4]])
```


<br>
<br>

## Principal Component Analysis (PCA) and clustering
The PCA is computed based on the coverage filtered methylation calls in the selected regions. 
```{r echo=FALSE}
clustering(data.cleanR,"UPN")
```

<br>
<br>

Let's furthermore look at the beta values of the regions. The following heatmap shows a hierachical clustering of the beta values in the selected regions. 
<br>

```{r echo=FALSE}
amp_heatmap(data.cleanR,"UPN",show_rownames=F)
```



